---
title: Creating a timeline using R
author: "Mauricio Vargas S. 帕夏"
Date: 2016-02-03
Category: R
Tags: R, rCharts, Python
Slug: timelinetutorial
Header_Cover: theme/images/cover.svg
Summary: Step by step tutorial to create a timeline that you can open on any browser.
output: html_document
---

```{r setup, cache = FALSE, echo = FALSE, message = FALSE, warning = FALSE, tidy = FALSE}
knitr::opts_chunk$set(message = F, error = F, warning = F, comment = NA, fig.align = 'center', dpi = 100, fig.width=6, fig.height=5, tidy = F, cache.path = '.cache/', fig.path = '../content/ggplottutorial2/')
```

On Jan, 10th David Bowie left this earthly realm. That day, as I do like his music a lot, I decided to create a [timeline](http://pachamaltese.github.io/bowie-timeline) and recently I decided to make the code publicly available.

## Required packages
You need [rCharts](http://ramnathv.github.io/rCharts/) whose author is [Ramnath Vaidyanathan](https://twitter.com/ramnath_vaidya). The package does not contain an extensive documentation yet but is a great tool.

## CSV file
Create a CSV file whose basic columns are

  * Start Date
  * End Date
  * Headline
  * Text
  * Media
  * Media Credit
  * Media Caption
  * Media Thumbnail
  * Type	
  * Tag

The only tricky part is that in Excel, Google Sheets or any other cellsheet software both start and end date must be formatted as m/d/yy hh:mm.

## Loading packages
```{r, eval=FALSE}
setwd("/Users/pacha/bowie-timeline/") #change this to your working folder
library(plyr)
library(RCurl)
library(rCharts)
```
If rChart is not installed it can be installed using devtools. The package website shows some deprecated parameters and this is what works.
```{r, eval=FALSE}
require(devtools)
install_github('ramnathv/rCharts')
```

## Reading data
```{r, eval=FALSE}
discography <- read.csv("discography.csv", as.is = TRUE)
discography <- alply(discography, 1, function(x) {
  list(startDate = x$Start.Date, headline = x$Headline, text = x$Text, asset = list(media = x$Media))
})
```

## Creating timeline
```{r, eval=FALSE}
tl = Timeline$new()
tl$main(headline = paste0("David Bowie"), type = "default", text = paste0("Discography ", 
        format(as.Date("1967-01-01"), "%Y"), "-", format(as.Date("2016-01-01"), "%Y")), 
        startDate = format(as.Date("1967-01-01"), "%Y"), asset = list(media = "img/cover.jpg"))
names(discography) <- NULL
tl$event(discography)

tl$save("index.html")
```

## Adjustments
Update: [Daisuke Ichikawa's blog](http://daisukeichikawa.blogspot.cl/2014/03/visualize-timeline-of-japanese-star.html) has a solution to change the JS path without editing the final html outside RStudio. This are the steps, do them or the timeline won't work in other computer.
```{r, eval=FALSE}
js <- paste(readLines("index.html", warn = F), collapse = "\n")
js <- gsub("/Users/pacha/bowie-timeline/timeline/js/storyjs-embed.js", 
          "timeline/js/storyjs-embed.js", 
          js)
writeLines(js, con = "index.html")
```

Now, if you open your timeline the dates will be a little bit ugly, look for `startDate` and the dates will be shown as "1/1/67 00:00", you can change it to "1967" without further steps but do the same to any `startDate` (e.g. use replace in your editor).

## Optional step
The final step is to go to the timeline folder that you just copied and get hands into the CSS file, then add the Creative Commons button and favicon, and then you will obtain exactly the same result as me. You can also add Google Analytics code to your timeline as with any html file.

## Verification
To avoid many commits to Github, you can run this from terminal 
```{r, eval=FALSE}
cd /Users/pacha/bowie-timeline
python -m SimpleHTTPServer
```
and then in your brower go to http://localhost:8000/ to see your changes.

## See the full project on Github
Go to the timeline repo available [here](https://github.com/pachamaltese/bowie-timeline).